Enabling Ambient Backscatter \\Using a Low-Cost Software Defined Radio

### 2017-04-07 paper.tex
Fixed figure 4.
A little cleanup in the results section

During the scanning process the gain is set to 40.2 dB. To get the
signal strength we simply take the average over a 10 MHz band around the
center frequency of the desired TV signal. This approach can be
justified by the fact, that DVB-T is specified up to 10 MHz bandwidth.
And the local TV signal provider is to its own statements using DVB-T 2.
The signal \ensuremath{|H|} is calculated as follows.  
\begin{equation}
	|H| = \Biggl| \sum_{n=0}^N FFT\biggl( \Re\{ s_{\text{band,n}}(t) \} \biggr) \Biggr|
\end{equation}     where \ensuremath{N} is the total number of intervals
accumulated over the entire frequency sweep. The FFT is only carried out over
the real values, which are received from the \textit{RTL-SDR}.
\ensuremath{s_{\text{band,n}}(t)} is the signal in the time domain of one
interval (one subband). There are as many samples taken in one band as needed
for an FFT of size 2048. To finally get \ensuremath{|H_{tv,avg}|} one has to
normalize everything and calculate the power. Before doing that a simple
smoothing algorithm (FIR) is applied on \ensuremath{|H|}.  
\begin{equation}
	|H_{\text{tv,avg}}| = 20 \cdot log (|H|) - 20 \cdot log(|H_0|)
\end{equation}   
where \ensuremath{|H_0|} is a normalization value measured at a certain point
in space. The result of combining measured signal strength values with the
haversine function to calculate the distance to the point where the maximum
signal has been captured can be observed in figure \ref{fig:haversine}.
\begin{figure}[h]
	\centering
	\begin{minipage}{0.49\columnwidth}
	\includegraphics[width=\columnwidth]{./fig/haversine}
	\end{minipage}
	\hfill
	\begin{minipage}{0.49\columnwidth}
	\includegraphics[width=\columnwidth]{./fig/heatmap_626mhz}
	\end{minipage}
	\caption{Left: Signal strength fading against distance relative
to the spot where the maximum signal strength has been captured (TV
towers). 474 MHz and 626 MHz belong to a tower in Uppsala, Vedyxa and
594 MHz belongs to a tower in Uppsala, Rickomberga. Distances have been
calculated with the Haversine formula. Right: Heatmap of the 626 MHz TV
signal in Uppsala.}

@@

\subsection{Communication Performance}
We tested the communication with different data rates from 1 bit/s up to
1 kbit/s. As one can see in figure \ref{fig:transmission} the signal
strength is quite low after backscattering. So the problem appears, that
not all bits of the A/D converter are used. Hence a lot of quantization
noise is introduced. In other words more gain is needed. The
transmission shown in figure \ref{fig:transmission} was carried out with
the maximum gain of 50 dB, which is available with the in
\ref{sub:rtl2832} decribed tuner. So we got to the limits of the
\textit{RTL-SDR} at that point. We were able to achieve a range indoors
with 1 bit/s of a couple of meteres, but when we use a higher bitrate we
can only communicate over a range of a couple of decimeter and still
have a bit error ratio of approximately 40 \%. The communiction
experiment was also carried out outside closer to the tower. We could
not figure out any improvements going closer to the tower.  
\section{Discussion}
We are aware of the fact, that our communication system has to be improved until it is able to be used in allday technology. These are aspects, where fine tuning is necessary to enhance the performance:
\begin{itemize}
	\item Antennas: Different backscattering antennas have to be
tried out to find the one with the best results. The question is mainly
to which frequency the antenna has to be tuned (e.g. to the center of
the TV signal or more to the edge of it). Moreover the standard
\textit{RTL-SDR} antenna including the coaxial cable are of quite low
quality. The antenna is e.g. too short. 
	\item Receiver: Some fine tuning can be made when it comes to the filter coefficients of the different filters.
	\item Frame: The frame structure can be optimized to be able to correct errors (e.g. Hamming code).  
\end{itemize}   
To come to a conclusion one can say, that we were able
\begin{itemize}
	\item to show, that spectrum scanning over a huge band with sweeping is possible using the \textit{RTL-SDR}.
	\item to provide a heat map of TV signals in a mid-sized swedish city as well as the tools to create it.
	\item to show for the first time, that backscatter communication is possible with the \textit{RTL-SDR}.
\end{itemize}

>>>>

During the scanning process the gain is set to 40.2 dB. To get the
signal strength we simply take the average over a 10 MHz band around the
center frequency of the desired TV signal. This approach can be
justified by the fact, that DVB-T is specified up to 10 MHz bandwidth.
And the local TV provider advertises to be transmitting DVB-T 2.
The signal \ensuremath{|H|} is calculated as follows.  
\begin{equation}
	|H| = \Biggl| \sum_{n=0}^N FFT\biggl( \Re\{ s_{\text{band,n}}(t) \} \biggr) \Biggr|
\end{equation}     where \ensuremath{N} is the total number of intervals
accumulated over the entire frequency sweep. The FFT is only carried out over
the real values, which are received from the \textit{RTL-SDR}.
\ensuremath{s_{\text{band,n}}(t)} is the signal in the time domain of one
interval (one subband). There are as many samples taken in one band as needed
for an FFT of size 2048. To finally get \ensuremath{|H_{tv,avg}|} one has to
normalize everything and calculate the power. Before doing that, a simple
smoothing algorithm (FIR) is applied on \ensuremath{|H|}.  
\begin{equation}
	|H_{\text{tv,avg}}| = 20 \cdot log (|H|) - 20 \cdot log(|H_0|)
\end{equation}   
where \ensuremath{|H_0|} is a normalization value measured at a certain point
in space. The result of combining measured signal strength values with the
haversine function to calculate the distance to the point where the maximum
signal has been captured can be observed in figure \ref{fig:haversine}.
\begin{figure}[h]
	\centering
	\begin{minipage}{0.49\columnwidth}
	\includegraphics[width=\columnwidth]{./fig/haversine}
	\end{minipage}
	\hfill
	\begin{minipage}{0.49\columnwidth}
	\includegraphics[width=\columnwidth]{./fig/heatmap_626mhz}
	\end{minipage}
	\caption{Left: Signal strength fading against distance relative
to the spot where the maximum signal strength has been captured (TV
towers). 474 MHz and 626 MHz belong to a tower in Uppsala, Vedyxa and
594 MHz belongs to a tower in Uppsala, Rickomberga. Distances have been
calculated with the Haversine formula. Right: Heatmap of the 626 MHz TV
signal in Uppsala.}

@@

\subsection{Communication Performance}
We tested the communication with different data rates from 1 bit/s up to
1 kbit/s. As can be seen in Figure \ref{fig:transmission} the signal
strength from the backscatter transmitter is rather low. This leads to a
significant amount of quantization noise to appear. This situation can
be improved by increasing the analog gain of the receiver. The
transmission shown in figure \ref{fig:transmission} was carried out with
the maximum gain (50 dB) available to our receiver.

With a data rate of 1 bit/s, we were able to achieve a range indoors
of a couple of meters. With higher bitrates we
can only communicate over a range of a few decimeters and the 
bit error rate is still approximately 40 \%. The communication
experiment was also carried out outside, close to the TV tower. We could
not observe any improvements going closer to the tower.

\section{Discussion}
We are aware of the fact, that our communication system has to be improved until it is able to be used in allday technology. These are aspects, where fine tuning is necessary to enhance the performance:
\begin{itemize}
	\item Antennas: Different backscattering antennas have to be
tried out to find the one with the best results. The question is mainly
to which frequency the antenna has to be tuned (e.g. to the center of
the TV signal or more to the edge of it). Moreover the standard
\textit{RTL-SDR} antenna including the coaxial cable are of quite low
quality. The antenna is e.g. too short. 
	\item Receiver: Some fine tuning can be made when it comes to the filter coefficients of the different filters.
	\item Frame: The frame structure can be optimized to be able to correct errors (e.g. Hamming code).  
\end{itemize}   
To come to a conclusion one can say, that we were able
\begin{itemize}
	\item to show, that spectrum scanning over a huge band with sweeping is possible using the \textit{RTL-SDR}.
	\item to provide a heat map of TV signals in a mid-sized swedish city as well as the tools to create it.
	\item to show for the first time, that backscatter communication is possible with the \textit{RTL-SDR}.
\end{itemize}
